---
title: "Billboard"
output:
  html_document:
    df_print: paged
---

```{r}
  library(tidyverse)
  library(billboard)
  library(ggthemes)
```

```{r}
  # Perform some tidying on the data before analysis.
  billboard_tidy <- billboard |>
    pivot_longer(
      cols = starts_with("wk"),
      names_to="week",
      values_to="rank",
      values_drop_na = TRUE,
    ) |>
    mutate(week=parse_number(week))
```

## Introduction

This notebook analyzes the **billboard** data set, which records the performance of songs on the top 100 charts for the year 2000.

## How long does a track typically remain in the top 100?

```{r}
  # Plot the time a song remained on the top 100 billboard in weeks, versus the
  # rank that the song achieved. This will give a good impression of how long that
  # a song remains on the top 100 in general.
  billboard_tidy |>
    ggplot(aes(x=week,y=rank,group=track)) +
    geom_line() +
    # Flip the scale of the y axis so that the rank is from highest to lowest.
    # This makes it easier to see the relationship between ranking and time on
    # the charts.
    scale_y_reverse() +
    labs(x="Weeks",y="Ranking",title="Weeks on the Top 100 vs. Ranking Achieved")
```

It is clear from the above plot that **very** few songs remain in the top 100 for more than a period of 20 weeks.

## What ten songs remained in the top ten for the longest number of weeks?

```{r}
  billboard_tidy |>
  filter(rank <= 10) |>
  group_by(track) |>
  summarize(weeks_in_top10=n()) |>
  arrange(desc(weeks_in_top10)) |>
  slice_head(n=10) |>
  ggplot(aes(y=reorder(track,weeks_in_top10),x=weeks_in_top10,color="Weeks",fill="Weeks")) +
  geom_col() +
  labs(x="Weeks in Top 10", y="Track Name") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  theme(legend.position="none")
```

There appears to have been a tie for the most popular song of 2000, with **Everything You Want** and **Breathe** having spent equal amounts of time in the top 10.

## Which ten artists had the most songs on the top 100?

```{r}
  billboard_tidy |>
  distinct(track, .keep_all=TRUE) |>
  group_by(artist) |>
  summarize(count=n()) |>
  arrange(desc(count)) |>
  slice_head(n=10) |>
  ggplot(aes(y=reorder(artist,count),x=count,color="Count",fill="Count")) +
  geom_col() +
  labs(x="Songs in Top 100",y="Artist",title="Artists with Songs in the Top 100") +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

**Jay-Z** was the most popular artist of the year 2000, with **5** songs charting on the top 100 alone.